---
title: "package_tutorial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This section is based on making your code reproducible. 
**D**on't
**R**epeat
**Y**ourself

# Work with some data in degrees F

- Convert to C

```{r}
airtemps <- c(212, 32, 64, 0, -20)

airtemp1 <- (airtemps[1]-32)*5/9
airtemp2 <- (airtemps[2]-32)*5/9
airtemp3 <- (airtemps[3]-32)*5/9
airtemp4 <- (airtemps[4]-32)*5/9
airtemp5 <- (airtemps[5]-32)*5/9
airtemp6 <- (airtemps[6]-32)*5/9

airtemp_F <- (airtemps - 32) * 5/9
```

# Write a function to do that

```{r}
fahr_to_celsius <- function(fahr) {
    celsius <- (fahr - 32) * 5/9
    return(celsius)
}

celsius_temps <- fahr_to_celsius(airtemps)
fahr_to_celsius()
head(celsius_temps)
```
If you don't specify the `return()` it will return the last value in the function.
You can specify the default input of the function 'fahr=100`.


# Write a function to convert it the other way
```{r}
celsius_to_fahr <- function(celsius) {
    fahr <- (celsius * 9/5) + 32
    return(fahr)
} 

fahr_temps <- celsius_to_fahr(celsius_temps)
fahr_temps == airtemps
```
Sometimes there are rounding errors associated with different versions of R. So watch out for that when testing.
Testing
1. Create an array.
2. Put into the function.
    - Check that each value is what is expected
    - Does it work without a value
3. Encode this as a test to make sure your function doesn't change.


# Show how to document functions

Define the function before you call it.
Same as importing library.

```{r}
#' Converts numeric temperature data from degrees Fahrenheit to degrees Celsius
#' 
#'@param fahr the value in degrees Fahrenheit to be converted
#'@return celsius the converted value in degrees Celsius
celsius_to_fahr <- function(celsius) {
    fahr <- (celsius * 9/5) + 32
    return(fahr)
} 
```

`#'` is the syntax that is associated with roxygen2 package for when you develop your package