---
title: "Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It's good to have an index page for your website.
The index.html needs to have a lower case `i` or the first letter needs to be lower case.

General directory structure:

Docs/

    - index.html
    
    - all other hmtl files
    
    - to find the other html pages type in [userame].github.io/[project_name]/[file_name].html
    
## Other files from this workshop
- [data reformatting practice](data_reformat.html)

-------------
## Tutorial Start

Load data directly from repository:

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r, warning = F, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```



```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = FALSE)
```

## **What we're going to be doing today**

* calculate annual escapement by species and region
    * split, combine need to create a Year variable
      * `seperate`
      * `group_by`
      * `summarize`
    * group_by, summarize, sum
    * filter
      * `filter`
* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

annual_esc

```{r}
annual_esc <- esc %>%
    separate(sampleDate, into = c("Year", "Month", "Day"), sep = "-") %>%
    group_by(Year, SASAP.Region, Species) %>%
    summarize(count = sum(DailyCount)) %>%
    # filter(Species != "Salmon")
    filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum"))
  
head(annual_esc)
```


# Generate some static plots

Things you have to give to ggplot:

    + data frame
    + mapping (aes, aka "aesthetics")
    + environment (give the parent frame)
    
Doesn't draw the geometry of the plot until you give it a geometry function

    + It will only do the mapping until then
    + Change the colors in the geom function
    
```{r}
ggplot(annual_esc, mapping = aes(x = Species, y = count)) +
    geom_col(fill = "blue")
```


# Generate an interactive map