---
title: "Geospatial Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE,message=FALSE}
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(leaflet)
library(scales)
```

vectors = polygons, points, lines

raster = grids

```{r}
ak_shp_sf <- read_sf("shapefiles/ak_regions_simp.shp")
```

* `read_sf` and `st_read` are almost identical 
* you can work with spatial objects like they are dataframes using sf
* gives you a sticky geometry column

```{r}
plot(ak_shp_sf)
```

To check the projection
```{r}
st_crs(ak_shp_sf)
```


Same as in cartopy...
```{r}
ak_shp_sf <- ak_shp_sf %>%
    st_transform(crs = 3338) # 3338: Alaska Albers Equidistant Area

st_crs(ak_shp_sf)

plot(ak_shp_sf)
```

## Spatial Joins

Bring in population data and plot across regions

```{r}
pop <- read.csv("shapefiles/alaska_population.csv")
```

If you want to do proper geospatial analysis you should do this:
* Remove in this will get rid of the lat/lon columns
```{r}
pop_sf <- st_as_sf(pop,
                   coords = c("lng", "lat"),
                   crs = 4326,
                   remove = F) %>%
    st_transform(crs = 3338)
```

```{r}
pop_joined_sf <- st_join(pop_sf, ak_shp_sf, join = st_within)

plot(pop_joined_sf["region"]) # default will plot every column
```

## Group by and summarise

Operate as if using a dataframe(we're tossing out the geometry/grouped point columns)

?group_by.sf
```{r}
pop_region <- pop_joined_sf %>%
    as.data.frame() %>%
    group_by(region) %>%
    summarise(total_pop = sum(population))

head(pop_region)
```

```{r}
ak_pop_sf <- left_join(ak_shp_sf, pop_region)

plot(ak_pop_sf["total_pop"])
```

You can group and summarise across geometries

```{r}
ak_mgmt <- ak_pop_sf %>%
    group_by(mgmt_area) %>%
    summarise(total_pop = sum(total_pop), do_union = FALSE) # do_union = to join the geometries

plot(ak_mgmt["total_pop"])
```

## How to write shapefiles

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_population.shp")
```

## Static Plots

Put aes in the geom function so that it's easier to understand the layering
    * You have to include the `data =` in geom function
```{r}
ggplot() + 
    geom_sf(data = ak_pop_sf, aes(fill = total_pop)) + 
    geom_sf(data = pop_sf, size = 0.5) +
    theme_bw() +
    labs(fill = "Total Population") +
    scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma)
```


## Interactive maps with Leaflet

Leaflet wants to project for you
    * Unproject your data
    * Then you have to give it a projection on the fly
```{r}
epsg3338 <- leaflet::leafletCRS(
    crsClass = "L.Proj.CRS",
    code = "EPSG:3338",
    proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
    resolutions = 2^(16:7))
```

1. Create your projection with leafletCRS()
2. Transform your data to the right projection
3. Create your color palette
4. Plot with Leaflet

```{r}
ak_crs <- st_transform(ak_pop_sf, crs = 4326)
pop_crs <- st_transform(pop_sf, crs = 4326)
```


```{r}
pal <- colorNumeric(palette = "Reds", domain = ak_crs$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
    addPolygons(data = ak_crs,
                fillColor = ~pal(total_pop),
                weight = 1,
                color = "black",
                label = ~region,
                fillOpacity = 1) %>%
    addCircleMarkers(data = pop_crs,
                     lat = ~lat,
                     lng = ~lng,
                     radius = 2,
                     fillColor = "gray",
                     fillOpacity = 1,
                     weight = .25,
                     color = "black",
                     label = ~city)

m
```

